---
title: "Data Establishment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(pheatmap)
library(gtsummary)
```

## Data Source

**DASS-42 Responses**

Our group applied the dataset from Kaggle Depression, Anxiety, and Stress Scales Responses (can be find [here]([https://www.kaggle.com/datasets/lucasgreenwell/depression-anxiety-stress-scales-responses](https://www.kaggle.com/datasets/lucasgreenwell/depression-anxiety-stress-scales-responses/data?select=codebook.txt))) for this project. The DASS-42 is a self-report scale consisting of 42 items intended to gauge emotional states related to depression, anxiety, and stress. The primary benefit of utilizing the DASS in a clinical context is its ability to ascertain the specific location of emotional turmoil, contributing to a more comprehensive clinical assessment. Its core function is to evaluate the severity of the fundamental symptoms associated with depression, anxiety, and stress.

```{r}
# Load the dataset
dass_raw = read.csv("./data/data.csv", sep = "\t")
```

The raw data contain `r nrow(dass_raw)` observations and `r ncol(dass_raw)` variables, where contains the information of the participants' responses.

## Data Establishment and Data Cleaning

As we mentioned before, the raw data contains the basic information from the responses. Based on codebook.txt from Kaggle, we checked whether the individual response is meaningful. The criteria are as follow:

-   The survey includes an accuracy check section, which provides several word spelling questions to check whether participants had filled out the questionnaire carefully. `VCL6`,`VCL9`,`VCL12` represents words that not exist. If the participants chose 1 in any of these three questions, the response could be omit.

-   The raw data contains different missing values which needs to be omitted. Specifically, `country` variable has "NONE" as the missing value.

-   `Age` variable have impossible values we need to clean out. For example, 299 years for an age is not reasonable.

```{r}
# kick out meaningless observation
dass = dass_raw %>%
  filter(!(VCL6 == 1 | VCL9 == 1 | VCL12 == 1)) %>%
  filter(!country == "NONE") %>%
  filter(age <= 100)

dass = dass %>%
  dplyr::select(matches('Q\\d+A'), country, education, urban, gender, age, religion, orientation, race, married) %>%
  na.omit()
```

Since we observed that there are some variables have actual meaning corresponding to the original numeric value, we refactor these variables, including `education`, `urban`, `gender`, `religion`, `orientation`, `race`, `married`. In these variables, if the value is 0, it represents missing value and we delete these observations.

```{r}
# refactor the variables
dass_new = dass %>%
  mutate(
    education = case_when(education == 1 ~ 'Less than high school', 
                               education == 2 ~ 'High school', 
                               education == 3 ~ 'University degree', 
                               education == 4 ~ 'Graduate degree'),
    urban = case_when(urban == 1 ~ 'Rural',
                      urban == 2 ~ 'Suburban', 
                      urban == 3 ~ 'Urban'),
    gender = case_when(gender == 1 ~ 'Male',
                       gender == 2 ~ 'Female',
                       gender == 3 ~ 'Other'),
    religion = case_when(religion == 1 ~ 'Agnostic', 
                         religion == 2 ~ 'Atheist', 
                         religion == 3 ~ 'Buddhist', 
                         religion == 4 ~ 'Christian (Catholic)', 
                         religion == 5 ~ 'Christian (Mormon)', 
                         religion == 6 ~ 'Christian (Protestant)', 
                         religion == 7 ~ 'Christian (Other)', 
                         religion == 8 ~ 'Hindu', 
                         religion == 9 ~ 'Jewish', 
                         religion == 10 ~ 'Muslim', 
                         religion == 11 ~ 'Sikh', 
                         religion == 12 ~ 'Other'),
    orientation = case_when(orientation == 1 ~ 'Heterosexual', 
                            orientation == 2 ~ 'Bisexual', 
                            orientation == 3 ~ 'Homosexual', 
                            orientation == 4 ~ 'Asexual', 
                            orientation == 5 ~ 'Other'),
    race = case_when(race == 10 ~ 'Asian', 
                     race == 20 ~ 'Arab', 
                     race == 30 ~ 'Black', 
                     race == 40 ~ 'Indigenous Australian', 
                     race == 50 ~ 'Native American', 
                     race == 60 ~ 'White', 
                     race == 70 ~ 'Other'),
    married = case_when(married == 1 ~ 'Never married', 
                        married == 2 ~ 'Currently married', 
                        married == 3 ~ 'Previously married')
    ) %>%
  na.omit()
```

Based on the scoring criteria, we add up the values for Q3A, Q5A, Q10A, Q13A, Q16A, Q17A, Q21A, Q24A, Q26A, Q31A, Q34A, Q37A, Q38A and Q42A as the depression scores. Similarly, add up the values for Q2A, Q4A, Q7A, Q9A, Q15A, Q19A, Q20A, Q23A, Q25A, Q28A, Q30A, Q36A, Q40A and Q41A as the anxiety scores and the sum of Q1A, Q6A, Q8A, Q11A, Q12A, Q14A, Q18A, Q22A, Q27A, Q29A, Q32A, Q33A, Q35A and Q39A as the stress scores. Considering that the scoring criteria we refer to is 0-3 scale but Kaggle dataset is 1-4 scale, we modify these three scores by minus 14.

After that, we could assign depression level, anxiety level and stress level for each observation refer to the rule of:

1.  Depression

    -   0 - 9 normal

    -   10 - 13 mild

    -   14 - 20 moderate

    -   21 - 27 severe

    -   \>27 extremely severe

2.  Anxiety

    -   0 - 7 normal

    -   8 - 9 mild

    -   10 - 14 moderate

    -   15 - 19 severe

    -   \>19 extremely severe

3.  Stress

    -   0 - 14 normal

    -   15 - 18 mild

    -   19 - 25 moderate

    -   26 - 33 severe

    -   \>33 extremely severe

Considering that stress, anxiety and depression might have correlations, we apply correlation test for `stress_level`, `depression_level`, `anxiety_level`.

```{r}
dass_heat2 <- dass %>%
  mutate(
    depression = Q3A + Q5A + Q10A + Q13A + Q16A + Q17A + Q21A + Q24A + Q26A + Q31A + Q34A + Q37A + Q38A + Q42A - 14,
    anxiety = Q2A + Q4A + Q7A + Q9A + Q15A + Q19A + Q20A + Q23A + Q25A + Q28A + Q30A + Q36A + Q40A + Q41A - 14,
    stress = Q1A + Q6A + Q8A + Q11A + Q12A + Q14A + Q18A + Q22A + Q27A + Q29A + Q32A + Q33A + Q35A + Q39A - 14
    ) %>%
  mutate(depression = case_when(
    depression >= 0 & depression <= 9 ~ 0,
    depression >= 10 & depression <= 13 ~ 1,
    depression >= 14 & depression <= 20 ~ 2,
    depression >= 21 & depression <= 27 ~ 3,
    depression >= 28 ~ 4
  )) %>%
  mutate(anxiety = case_when(
    anxiety >= 0 & anxiety <= 7 ~ 0,
    anxiety >= 8 & anxiety <= 9 ~ 1,
    anxiety >= 10 & anxiety <= 14 ~ 2,
    anxiety >= 15 & anxiety <= 19 ~ 3,
    anxiety >= 20 ~ 4
  )) %>%
  mutate(stress = case_when(
    stress >= 0 & stress <= 14 ~ 0,
    stress >= 15 & stress <= 18 ~ 1,
    stress >= 19 & stress <= 25 ~ 2,
    stress >= 26 & stress <= 33 ~ 3,
    stress >= 34 ~ 4
  )) %>%
  dplyr::select(depression,anxiety,stress)

r1 <- cor(dass_heat2,
         method = "pearson",
         use = "pairwise.complete.obs"
) 

pheatmap(r1, 
         show_colnames = TRUE,   
         show_rownames=TRUE,    
         fontsize=5,             
         color = colorRampPalette(c('#ffffff','#ff0000'))(100), 
         annotation_legend=TRUE, 
         border_color=NA,        
         scale="none",            
         cluster_rows = F,
         cluster_cols = F,
         breaks = seq(0, 1, length.out = 90)
)
```

According to the heat map, we could conclude that these three variables have strong correlation. Therefore, in our project, we take depression_level as the representative.

Finally, we get our clean and tidy dataset for the project.

```{r}
# create three different dataset
dass_new = dass_new %>%
  mutate(
    depression = Q3A + Q5A + Q10A + Q13A + Q16A + Q17A + Q21A + Q24A + Q26A + Q31A + Q34A + Q37A + Q38A + Q42A - 14
    ) %>%
  mutate(education = factor(education, levels = c('Less than high school', 'High school', 'University degree', 'Graduate degree'))) %>%
  mutate(married = factor(married, levels = c("Previously married", "Currently married", "Never married"))) %>%
  dplyr::select(-matches('Q\\d+A'))

# create das level variables
dass_new = dass_new %>%
  mutate(depression_level = case_when(
    depression >= 0 & depression <= 9 ~ 'Normal',
    depression >= 10 & depression <= 13 ~ 'Mild',
    depression >= 14 & depression <= 20 ~ 'Moderate',
    depression >= 21 & depression <= 27 ~ 'Severe',
    depression >= 28 ~ 'Extremely severe'
  )) %>%
  mutate(depression_level = factor(depression_level, c("Normal","Mild","Moderate","Severe","Extremely severe"))) %>%
  dplyr::select(-depression)

head(dass_new)
```

## Data Summary

Above all, we got our dataset for further analysis. It contains `r nrow(dass_new)` observations and `r ncol(dass_new)` variables. The explanation of each variables are listed below:

-   `country`: The country where the participants from.

-   `education`: The education level of the participants.

-   `urban`: The urban level of the place where participants live.

-   `gender`: The sex of the participants.

-   `religion`: The religion which the participants belong to.

-   `orientation`: The sexual orientation of the participants.

-   `race`: The race of participants.

-   `married`: The marriage status of participants.

-   `age`: The age when participants response the survey.

-   `depression_level`: The depression level of the participants according to the survey and scores.

```{r}



dass_new %>%
  dplyr::select(-country) %>%
  tbl_summary(type = list(
    education ~ "categorical", 
    urban ~ "categorical", 
    gender ~ "categorical",
    religion ~ "categorical",
    orientation ~ "categorical",
    race ~ "categorical",
    married ~ "categorical",
    age ~ "continuous", 
    depression_level ~ "categorical"
  ), 
  statistic = list(all_continuous() ~ "{mean} ({sd})"), 
  digits = all_continuous() ~ 1,
  label = c(
    education = "Education level",
    urban = "Urban level",
    gender = "Gender",
    religion = "Religion",
    orientation = 'Sexual orientation',
    race = 'Race',
    married = 'Marriage status',
    age = "Age (Years)",
    depression_level = "Depression level"
  )) %>%
  bold_labels() %>%
  as_gt()  # Display the table as a gt object

```
