---
title: "Exploratory Data Analysis"
output: html_document
---

```{r, setup, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(ggridges)
dass_new = read.csv("./data/dass_new.csv")
```

Shiny app list here

# Plot:
## Depression Level vs. Orientation vs. Gender
```{r}
dass_new|>
  group_by(gender, orientation)|>
  summarise(Normal = sum(depression_level=="Normal"),
          Mild = sum(depression_level=="Mild"),
          Moderate = sum(depression_level=="Moderate"),
          Severe = sum(depression_level=="Severe"),
          Extremely_Severe = sum(depression_level=="Extremely severe"))|>
  pivot_longer(
    Normal:Extremely_Severe,
    names_to = "Level",
    values_to = "count")|>
  mutate(Level = factor(Level, levels = unique(Level)))|>
  ggplot(aes(x = Level, y = count, fill= gender)) +
  geom_col()+
  theme_minimal()+
  facet_grid(~orientation)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Depression level", 
       y = "count of cases", 
       title = "Depression level vs. orientation vs. gender")
```

## Depression level vs. Orientation

```{r}
dass_new|>
  group_by(orientation)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))|>
  pivot_longer(
    Normal:Extremely_Severe,
    names_to = "Level",
    values_to = "Count")|>
  mutate(Level = factor(Level, levels = unique(Level)))|>
  ggplot(aes(x = Level, y = Count, group = orientation, color = orientation)) +
  geom_point()+
  geom_line()+
  theme_minimal()+
  labs(x = "Depression level", 
       y = "count of cases", 
       title = "Depression level vs. orientation groups")
```

## Depression level vs. Gender

```{r}
# distribution histogram of gender
dass_new|>
  group_by(gender)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))|>
  pivot_longer(
    Normal:Extremely_Severe,
    names_to = "Level",
    values_to = "Count")|>
  mutate(Level = factor(Level, levels = unique(Level)))|>
  ggplot(aes(x = Level, y = Count, group = gender, color = gender)) +
  geom_point()+
  geom_line()+
  theme_minimal()+
  labs(x = "Depression level", 
       y = "count of cases", 
       title = "Depression level vs. gender groups")
```


## Depression Level vs. Race groups

```{r}
dass_new|>
  # form new race groups
  mutate(race = case_when(race == "Native American" ~ "Other",
                          race=="Arab"~ "Other", 
                          race == "Indigenous Australian" ~ "Other",
                          race == "Other" ~ "Other",
                          race == "Black" ~ "Black",
                          race == "White" ~ "White",
                          race == "Asian" ~ "Asian"))|>
  group_by(race)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))|>
  pivot_longer(
    Normal:Extremely_Severe,
    names_to = "Level",
    values_to = "Count")|>
  mutate(Level = factor(Level, levels = unique(Level)))|>
  ggplot(aes(x = Level, y = Count, group = race, color = race)) +
  geom_point()+
  geom_line()+
  theme_minimal()+
  labs(x = "Depression level", 
       y = "count of cases", 
       title = "Depression level vs. race groups")
```

## Depression Level vs. Age groups

```{r}
dass_new|>
  # form age groups
  mutate(age = case_when(age <= 18 ~ "teenage", 
                         age <= 25 & age >18 ~ "young adult",
                         age <= 50 & age >25 ~ "adult", 
                         age >50 ~ "older"))|>
  group_by(age)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))|>
  pivot_longer(
    Normal:Extremely_Severe,
    names_to = "Level",
    values_to = "Count")|>
  mutate(Level = factor(Level, levels = unique(Level)))|>
  ggplot(aes(x = Level, y = Count, group = age, color = age)) +
  geom_line()+
  geom_point()+
  theme_minimal()+
  labs(x = "Depression level", 
       y = "count of cases", 
       title = "Depression level vs. age groups")
```

## Age Distribution for Each Depression Level
```{r}
dass_new %>%
  ggplot(aes(x=depression_level, y=age)) +
  geom_boxplot(aes(fill=depression_level)) +
  labs(title = "Age Distribution for Each Depression Level",
        xlab = "Depression Level",
        ylab = "Age")+
  theme_minimal()

```


# Chi-squared Test

## 1: Orientation & Depression Level

From the plot of depression level vs. orientation, we observed that the distribution of cases of different depression levels among different orientation groups follow a similar pattern. Thus, we would like to perform a chi-squared test to study if the depression levels and orientation are associated.\

$H_0$: The depression levels and orientation are independent.\
$H_1$: The depression levels and orientation are associated.

```{r}
# show result table to perform chi-squared test
dass_new|>
  group_by(orientation)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))
# perform chi-squared test
chisq.test(pull(dass_new, orientation), pull(dass_new, depression_level))|>
  broom::tidy()|>
  knitr::kable()
```

The null hypothesis for the chi-squared test is that the depression levels and orientation are independent. Since the p-value is smaller than 0.05, the null hypothesis is rejected. \
We may conclude that orientation is associated with depression levels.


## 2: Gender & Depression Level

From the plot of depression level vs. gender, we observed that the distribution of cases of different depression levels among different gender groups follow a similar pattern. Thus, we would like to perform a chi-squared test to study if the depression levels and gender are associated.\

$H_0$: The depression levels and gender are independent.\
$H_1$: The depression levels and gender are associated.

```{r}
# show result table to perform chi-squared test
dass_new|>
  group_by(gender)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))
# perform chi-squared test
chisq.test(pull(dass_new, gender), pull(dass_new, depression_level))|>
  broom::tidy()|>
  knitr::kable()
```

The null hypothesis for the chi-squared test is that the depression levels and gender are independent. Since the p-value is smaller than 0.05, the null hypothesis is rejected. \
We may conclude that gender is associated with depression levels.


## 3: Age & Depression Level

From the plot of depression level vs. age groups, we observed that the distribution of cases of different depression levels among different age groups follow a similar pattern. Thus, we would like to perform a chi-squared test to study if the depression levels and age are associated.\

$H_0$: The depression levels and age are independent.\
$H_1$: The depression levels and age are associated.

```{r}
dass_chi <- dass_new|>
  # form age groups
  mutate(age = case_when(age <= 18 ~ "teenage", 
                         age <= 25 & age >18 ~ "young adult",
                         age <= 50 & age >25 ~ "adult", 
                         age >50 ~ "older"))
# show result table to perform chi-squared test
dass_chi|>
  group_by(age)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))
# perform chi-squared test
chisq.test(pull(dass_chi, age), pull(dass_chi, depression_level))|>
  broom::tidy()|>
  knitr::kable()
```

The null hypothesis for the chi-squared test is that depression levels and age are independent. Since the p-value is smaller than 0.05, the null hypothesis is rejected.\
We may conclude that age is associated with depression levels.

## 4: Race & Depression Level

Similarly, we observed that the distribution of cases of different depression levels among different race groups follow a similar pattern from the plot of depression levels vs. race groups. Thus, we would like to perform a chi-squared test to study if the depression levels and race groups are associated.\

$H_0$: The depression levels and race groups are independent.\
$H_1$: The depression levels and race groups are associated.

```{r}
dass_chi <- dass_new|>
  # form new race groups
  mutate(race = case_when(race == "Native American" ~ "Other",
                          race=="Arab"~ "Other", 
                          race == "Indigenous Australian" ~ "Other",
                          race == "Other" ~ "Other",
                          race == "Black" ~ "Black",
                          race == "White" ~ "White",
                          race == "Asian" ~ "Asian"))
# show result table to perform chi-squared test
dass_chi|>
  group_by(race)|>
  summarise(Normal = sum(depression_level=="Normal"),
            Mild = sum(depression_level=="Mild"),
            Moderate = sum(depression_level=="Moderate"),
            Severe = sum(depression_level=="Severe"),
            Extremely_Severe = sum(depression_level=="Extremely severe"))
# perform chi-squared test
chisq.test(pull(dass_chi, race), pull(dass_chi, depression_level))|>
  broom::tidy()|>
  knitr::kable()
```

The null hypothesis for the chi-squared test is that depression levels and race groups are independent. Since the p-value is smaller than 0.05, the null hypothesis is rejected. \
We may conclude that race is associated with depression levels.

