---
title: "Logistic Model"
author: "Shaolei Ma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)
library(car)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

The goal is to fit a ordered logit model.

### Independent Variables

First, test the distribution of independent variables.
```{r}
dass_new = read_csv("data/dass_new.csv")

dass_new |> 
  mutate(
    country = as.factor(country),
    urban = as.factor(urban),
    gender = as.factor(gender),
    religion = as.factor(religion),
    orientation = as.factor(orientation),
    race = as.factor(race)
  ) |> 
  summary()
```

Some categories contain few samples. For the efficacy of the model, we regroup the independent variables following the standards below:

 * `country`: Group the country with less than 500 samples into "Other"
 * `race`: We follow [standards on race and ethnicity](https://www.govinfo.gov/content/pkg/FR-1997-10-30/pdf/97-28653.pdf) to include only "Asian", "Black", "Native American", "White", and make other categories "Other".
 * `religion`: We conclude "Christian (Catholic)", "Christian (Mormon)", "Christian (Other)", "Christian (Protestant)" into "Christian".
 * `orientation`: Set the "Heterosexual" to be the reference level.

```{r}
dass_fit_df =
  dass_new |> 
  mutate(
    country = as.factor(ifelse(country %in% c("MY", "US", "GB", "CA", "ID", "PH"), country, "Other")),
    race = as.factor(ifelse(race %in% c("Arab", "Indigenous Australian"), "Other", race)),
    religion = as.factor(ifelse(substr(religion, 1, 9) == "Christian", "Christian", religion)),
    urban = as.factor(urban),
    gender = as.factor(gender),
    orientation = relevel(as.factor(orientation), "Heterosexual"),
    depression_level = factor(depression_level, levels = c("Normal", "Mild", "Moderate", "Severe", "Extremely severe"))
  )

dass_fit_df |> summary()
```

### Model Fitting

First fit a model without variables selection.
```{r}
dass_fit = 
  dass_fit_df |> 
  polr(depression_level ~ ., data = _)

(fit_s = 
  dass_fit |> 
  summary())
```

The following is the significant variables:
```{r echo=FALSE}
fit_s$coefficients[,1][abs(fit_s$coefficients[,3]) < 1.96] |> knitr::kable()
```

It could be observed that all dummy variables belonging to `education`, `gender`, `orientation`, `married` are not significant, and the continuous variable `sex` is not significant. We exclude these three variables and refit the model.

```{r}
dass_fit2 = 
  dass_fit_df |> 
  polr(depression_level ~ country + urban + religion + race, data = _)

dass_fit2 |> 
  summary()
```

Test the parallel assumption.
```{r}
car::poTest(dass_fit2)
```

Because the overall p value < 0.001, we reject the null hypothesis and conclude that the parallel assumption is not met.

Therefore, we regroup the dependent variable into "Moderate or Below", "Severe", "Extremely severe" and refit the model.

```{r}
dass_fit3 =
  dass_fit_df |> 
  mutate(depression_level = factor(case_match(
    depression_level,
    "Normal" ~ "Moderate or Below",
    "Mild" ~ "Moderate or Below","Moderate" ~ "Moderate or Below",
    "Severe" ~ "Severe",
    "Extremely severe" ~ "Extremely severe"
   ), level = c("Moderate or Below", "Severe", "Extremely severe"))) |> 
  polr(depression_level ~ country + urban + religion + race, data = _)

dass_fit3 |>  summary()

dass_fit3 |> poTest()
```

The model meets the parallel assumption.

### Future Work

 * Re-selection of independent variables. (Significant, Regroup, AIC, PCA)
 * Interpretation of OR.