---
title: "Logistic Model"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(foreign)
library(ggplot2)
library(MASS)
library(Hmisc)
library(reshape2)
library(car)
library(pROC)
library(effects)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%",
  message = F,
  warning = F
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Now, we want to investigate the relationship between demographic variables and depression levels. Because the dependent variable `depression_level` is an ordinal variable with five levels ("Normal", "Mild", "Moderate", "Severe", "Extremely severe"), we choose to fit an ordered logit model.

# Independent Variables

Imbalanced data may result in an inaccurate outcome due to a small sample size in some classes. Therefore, we first want to test the distribution of independent variables to avoid the imbalance of samples between different categories.

```{r}
dass_new = read_csv("data/dass_new.csv")

dass_new |> 
  dplyr::select(age, everything()) |> 
  mutate(across(country:depression_level, factor)) |> 
  summary()
```

Some categories contain few samples (e.g. countries other than the listed six countries have less than 500 samples). For the efficacy of the model and considering which class most interests us, we regroup the independent variables following the standards below:

-   `country`: Group the country with less than 500 samples into "Other", and set "US" to be the reference level.
-   `education`: Order the variable by the length of education.
-   `race`: We follow [standards on race and ethnicity](https://www.govinfo.gov/content/pkg/FR-1997-10-30/pdf/97-28653.pdf) to include only "Asian", "Black", "Native American", "White", and make other categories "Other". We set "White" to be the reference level.
-   `religion`: We conclude "Christian (Catholic)", "Christian (Mormon)", "Christian (Other)", "Christian (Protestant)" into "Christian", and set it as the reference level.
-   `orientation`: Set the "Heterosexual" to be the reference level.
-   `married`: Order the variable by marriage status.

```{r}
dass_fit_df =
  dass_new |> 
  mutate(
    country = relevel(as.factor(ifelse(country %in% c("MY", "US", "GB", "CA", "ID", "PH"), country, "Other")), ref = "US"),
    education = factor(education, levels = c("Less than high school", "High school", "University degree", "Graduate degree")),
    race = relevel(as.factor(ifelse(race %in% c("Arab", "Indigenous Australian"), "Other", race)), ref = "White"),
    religion = relevel(as.factor(ifelse(substr(religion, 1, 9) == "Christian", "Christian", religion)), "Christian"),
    orientation = relevel(as.factor(orientation), "Heterosexual"),
    married = factor(married, levels = c("Never married", "Currently married", "Previously married")),
    depression_level = factor(depression_level, levels = c("Normal", "Mild", "Moderate", "Severe", "Extremely severe"))
  )

dass_fit_df |> summary()
```

# Model Fitting

First fit a model without variables selection.

```{r}
dass_fit = 
  dass_fit_df |> 
  polr(depression_level ~ ., data = _)

(fit_s = 
  dass_fit |> 
  summary())
```

When conducting an ordered logit model, the underlying assumption is that the coefficients that describe the relationship between, say, the lowest versus all higher categories of the response variable are the same as those that describe the relationship between the next lowest category and all higher categories, etc. This is called the proportional odds assumption or the parallel assumption.

Now, we want to test if the model conducted meets the parallel assumption.

```{r}
car::poTest(dass_fit)
```

Because the overall $p-value < 0.001$, we reject the null hypothesis and conclude that the parallel assumption is not met.

To meet the parallel assumption, considering the variables of interest in EDA, we try the following two methods:

-   **Select independent variables**: Because we think `religion` and `race`, `married` and `orientation`, `country` and `urban` are strongly correlated with each other, and we care more about `race`, `orientation`, and `country`, first delete `religion`, `married`, and `urban`. We also delete `education` to meet the parallel assumption.
-   **Regroup dependent variable**: Regroup `depression_level` into "Moderate or Below", "Severe", "Extremely severe".

```{r}
dass_fit_df =
  dass_fit_df |> 
  mutate(depression_level = factor(case_match(
    depression_level,
    "Normal" ~ "Moderate or Below",
    "Mild" ~ "Moderate or Below","Moderate" ~ "Moderate or Below",
    "Severe" ~ "Severe",
    "Extremely severe" ~ "Extremely severe"
   ), level = c("Moderate or Below", "Severe", "Extremely severe")))

dass_fit1 =
  dass_fit_df |> 
  polr(depression_level ~ country + gender + age + orientation + race, data = _)

dass_fit1 |> poTest()
```

The model meets the parallel assumption at $\alpha=0.05$ level (with an overall $p = 0.067 > 0.05$).

## Depression Levels Distribution

After selecting independent variables, we can further visualize the data based on the effective covariates as well as the good stratum of depression level.

```{r, fig.width=20, fig.height=50, message = FALSE}
# Create a data frame with the predictor values for visualization
predictor_data <- model.matrix(depression_level ~ country + gender + age
                               + orientation + race, data = dass_fit_df)

# Compute the effects
effects <- allEffects(dass_fit1, xlevels = list(predictor_data))

# Plot the effects
plot(effects, style = "stacked")
```

The accompanying graphs display the probability of depression levels across various categorical and continuous variables. For instance, the trend depicted in the graph indicates higher rates of extreme depression among young people.

# Results

```{r}
dass_fit1 |> summary()
```

The estimated model can be written as: \begin{align*}
\text{logit}(\hat P(\text{Depression Level} \leq \text{Severe})) &= -0.69 \\
&+ 0.18 \times I(\text{country} = \text{CA}) \\
&+ 0.20 \times I(\text{country} = \text{GB}) \\
&- 0.04 \times I(\text{country} = \text{ID}) \\
&- 0.19 \times I(\text{country} = \text{MY}) \\
&+ 0.13 \times I(\text{country} = \text{PH}) \\
&+ 0.11 \times I(\text{country} = \text{Other}) \\
&- 0.10 \times I(\text{gender} = \text{Male}) \\
&+ 0.38 \times I(\text{gender} = \text{Other}) \\
&- 0.03 \times \text{age} \\
&+ 0.23 \times I(\text{orientation} = \text{Asexual}) \\
&+ 0.40 \times I(\text{orientation} = \text{Bisexual}) \\
&+ 0.30 \times I(\text{orientation} = \text{Homosexual}) \\
&+ 0.22 \times I(\text{orientation} = \text{Other}) \\
&- 0.16 \times I(\text{race} = \text{Asian}) \\
&- 0.09 \times I(\text{race} = \text{Black}) \\
&+ 0.11 \times I(\text{race} = \text{Native American}) \\
&- 0.10 \times I(\text{race} = \text{Other})
\end{align*}

```{=tex}
\begin{align*}
\text{logit}(\hat P(\text{Depression Level} \leq \text{Extremely Severe})) &= 0.01 \\
&+ 0.18 \times I(\text{country} = \text{CA}) \\
&+ 0.20 \times I(\text{country} = \text{GB}) \\
&- 0.04 \times I(\text{country} = \text{ID}) \\
&- 0.19 \times I(\text{country} = \text{MY}) \\
&+ 0.13 \times I(\text{country} = \text{PH}) \\
&+ 0.11 \times I(\text{country} = \text{Other}) \\
&- 0.10 \times I(\text{gender} = \text{Male}) \\
&+ 0.38 \times I(\text{gender} = \text{Other}) \\
&- 0.03 \times \text{age} \\
&+ 0.23 \times I(\text{orientation} = \text{Asexual}) \\
&+ 0.40 \times I(\text{orientation} = \text{Bisexual}) \\
&+ 0.30 \times I(\text{orientation} = \text{Homosexual}) \\
&+ 0.22 \times I(\text{orientation} = \text{Other}) \\
&- 0.16 \times I(\text{race} = \text{Asian}) \\
&- 0.09 \times I(\text{race} = \text{Black}) \\
&+ 0.11 \times I(\text{race} = \text{Native American}) \\
&- 0.10 \times I(\text{race} = \text{Other})
\end{align*}
```
Calculate the odds ratio and CI for variables:

```{r}
exp(cbind(OR = coef(dass_fit1), confint(dass_fit1))) |> 
  knitr::kable()
```

## Receiver Operating Characteristic Curve for Depression Level

```{r}

# Predicting probabilities for all classes
predicted_probs <- predict(dass_fit1, type = "probs")

# Extracting the response variable
response <- ifelse(dass_fit_df$depression_level == "Moderate or Below", 1,
                   ifelse(dass_fit_df$depression_level == "Severe", 2, 3))

# Calculating ROC curves for three responses
roc_obj_1 <- roc(as.numeric(response == 1), predicted_probs[, "Moderate or Below"])
roc_obj_2 <- roc(as.numeric(response == 2), predicted_probs[, "Severe"])
roc_obj_3 <- roc(as.numeric(response == 3), predicted_probs[, "Extremely severe"])

# Plotting ROC curves for three responses on the same graph
plot(roc_obj_1, col = "blue", main = "ROC Curves for Three Responses")
plot(roc_obj_2, col = "red", add = TRUE)
plot(roc_obj_3, col = "green", add = TRUE)
legend("bottomright",
       legend = c("Moderate or Below", "Severe", "Extremely Severe"),
       col = c("blue", "red", "green"), lty = 1)
```

Looking at the ROC curves, it is clear that both the "Moderate or Below" and "Extremely Severe" categories show good performance and are close to the upper left corner, indicating effective performance in discriminating between these categories. However, the performance of the "Severe" category suggests an available but less significant trend compared to the above two categories. The positioning of this curve indicates room for improvement, implying a need for more data points representing individuals with a "severe" level of depression. Expanding the data set to include additional instances characterized by a "Severe" level of depression could improve the model's ability to discriminate and accurately predict within this specific category.

# Interpretation

## Continuous Variable

### **Age**

-   For every one unit increase in participant's age, the odds of having a more severe depression level is multiplied 0.98 times, holding constant all other variables. It indicates that older people may have less risk to get depressed.

## Categorical Variables

### **Country**

```{r}
# Extract coefficients for countries
pos_country = str_starts(names(dass_fit1$coefficients), "country")

country_coefficients =
  tibble(
    country = substr(names(dass_fit1$coefficients)[pos_country], 8, 20),
    OR = exp(dass_fit1$coefficients[pos_country])
  ) |> 
  mutate(country = fct_reorder(country, OR))
  

# Plotting coefficients for countries
ggplot(country_coefficients, aes(x = country, y = OR - 1)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Effect of Countries on Depression Levels",
       x = "Country", y = "OR - 1")
```

-   For Canadian, Britain, and Philippines participants, the odds of having a more severe depression level is higher than that of US participants, holding constant all other variables.
-   For Indonesia and Malaysia participants, the odds of having a more severe depression level is lower than that of US participants, holding constant all other variables.

### **Gender**

```{r}
# Extract coefficients for Gender
pos_gender = str_starts(names(dass_fit1$coefficients), "gender")

gender_coefficients =
  tibble(
    gender = substr(names(dass_fit1$coefficients)[pos_gender], 7, 20),
    OR = exp(dass_fit1$coefficients[pos_gender])
  ) |> 
  mutate(gender = fct_reorder(gender, OR))

# Plotting coefficients for Gender
ggplot(gender_coefficients, aes(x = gender, y = OR - 1)) +
  geom_bar(stat = "identity", fill = "#d8b4e2", color = "black") +
  labs(title = "Effect of Gender on Depression Levels",
       x = "Gender", y = "OR - 1")
```

-   For male participants, the odds of having a more severe depression level is 0.90 times that of female participants, holding constant all other variables.
-   For non-binary participants, the odds of having a more severe depression level is 1.47 times that of female participants, holding constant all other variables.

### **Race**

```{r}
# Extract coefficients for races
pos_race <- str_starts(names(dass_fit1$coefficients), "race")

race_coefficients =
  tibble(
    race = substr(names(dass_fit1$coefficients)[pos_race], 5, 20),
    OR = dass_fit1$coefficients[pos_race]
  ) |> 
  mutate(race = fct_reorder(race, OR))

# Plotting coefficients for races
ggplot(race_coefficients, aes(x = race, y = OR - 1)) +
  geom_bar(stat = "identity", fill = "pink", color = "black") +
  labs(
    title = "Effect of Races on Depression Levels",
    x = "Race",
    y = "OR - 1"
  )
```

-   For Native American, Asian, Black, and Other participants, the odds of having a more severe depression level is lower than that of White participants, holding constant all other variables.

### **Orientation**

```{r}
# Extract coefficients for orientations
pos_orientation <- str_starts(names(dass_fit1$coefficients), "orientation")

orientation_coefficients <- data.frame(
  orientation = substr(names(dass_fit1$coefficients)[pos_orientation], 12, 20),
  OR = exp(dass_fit1$coefficients[pos_orientation])
)

# Plotting coefficients for orientations
ggplot(orientation_coefficients, aes(x = orientation, y = OR - 1)) +
  geom_bar(stat = "identity", fill = "palegreen", color = "black") +
  labs(
    title = "Effect of Orientations on Depression Levels",
    x = "Orientation",
    y = "OR - 1"
  )
```

-   For sexual minorities, the odds of having a more severe depression level is higher than that of heterosexual orientation participants, holding constant all other variables.

# Limitations

-   Imbalance within samples (e.g. 76.3% female samples).
-   Strong dependence between independent variables (e.g. race versus religion).
-   Counter-intuitive results (e.g. Asian people have lower risk of getting severely depressed than white people).
-   Future Improvements:
    -   Clustering to decrease dependence;
    -   Fit a separate model for each depression level to avoid parallel assumption.
