---
title: "Logistic Model"
author: "Huanyu Chen, Shaolei Ma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)
library(car)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

The goal is to fit an ordered logit model.

## Independent Variables

First, test the distribution of independent variables.
```{r}
dass_new = read_csv("data/dass_new.csv")

dass_new |> 
  mutate(
    country = as.factor(country),
    urban = as.factor(urban),
    gender = as.factor(gender),
    religion = as.factor(religion),
    orientation = as.factor(orientation),
    race = as.factor(race)
  ) |> 
  summary()
```

Some categories contain few samples. For the efficacy of the model, we regroup the independent variables following the standards below:

 * `country`: Group the country with less than 500 samples into "Other"
 * `race`: We follow [standards on race and ethnicity](https://www.govinfo.gov/content/pkg/FR-1997-10-30/pdf/97-28653.pdf) to include only "Asian", "Black", "Native American", "White", and make other categories "Other".
 * `religion`: We conclude "Christian (Catholic)", "Christian (Mormon)", "Christian (Other)", "Christian (Protestant)" into "Christian".
 * `orientation`: Set the "Heterosexual" to be the reference level.

```{r}
dass_fit_df =
  dass_new |> 
  mutate(
    country = as.factor(ifelse(country %in% c("MY", "US", "GB", "CA", "ID", "PH"), country, "Other")),
    race = as.factor(ifelse(race %in% c("Arab", "Indigenous Australian"), "Other", race)),
    religion = as.factor(ifelse(substr(religion, 1, 9) == "Christian", "Christian", religion)),
    urban = as.factor(urban),
    gender = as.factor(gender),
    orientation = relevel(as.factor(orientation), "Heterosexual"),
    depression_level = factor(depression_level, levels = c("Normal", "Mild", "Moderate", "Severe", "Extremely severe"))
  )

dass_fit_df |> summary()
```

## Model Fitting

First fit a model without variables selection.
```{r}
dass_fit = 
  dass_fit_df |> 
  polr(depression_level ~ ., data = _)

(fit_s = 
  dass_fit |> 
  summary())
```

The following is the significant variables:
```{r echo=FALSE}
fit_s$coefficients[,1][abs(fit_s$coefficients[,3]) < 1.96] |> knitr::kable()
```

It could be observed that all dummy variables belonging to `education`, `gender`, `orientation`, `married` are not significant, and the continuous variable `sex` is not significant. We exclude these three variables and refit the model.

```{r}
dass_fit2 = 
  dass_fit_df |> 
  polr(depression_level ~ country + urban + religion + race, data = _)

dass_fit2 |> 
  summary()
```

Test the parallel assumption.
```{r}
car::poTest(dass_fit2)
```

Because the overall p value < 0.001, we reject the null hypothesis and conclude that the parallel assumption is not met.

Therefore, we regroup the dependent variable into "Moderate or Below", "Severe", "Extremely severe" and refit the model.

```{r}
dass_fit3 =
  dass_fit_df |> 
  mutate(depression_level = factor(case_match(
    depression_level,
    "Normal" ~ "Moderate or Below",
    "Mild" ~ "Moderate or Below","Moderate" ~ "Moderate or Below",
    "Severe" ~ "Severe",
    "Extremely severe" ~ "Extremely severe"
   ), level = c("Moderate or Below", "Severe", "Extremely severe"))) |> 
  polr(depression_level ~ country + urban + religion + race, data = _)

dass_fit3 |>  summary()

dass_fit3 |> poTest()
```

Generally speaking, the model meets the parallel assumption (with an overall p = 0.121). But there exist several individual predictors that do not meet parallel assumption (e.g., `urbanSuburban` with p = 0.049 and `religionAtheist` with p = 0.026). Since the `country` variable itself contains information about region and area, we want to remove the group for urban and suburban and run the parallel assumption again.

Utilizing PCA to test if `urban` variable can be deleted:
```{r}
selected_data <- dass_fit_df[, c("education", "urban", "age", "religion", "race")]
scaled_encoded_data <- scale(model.matrix(~ . - 1, data = selected_data))
pca_result <- prcomp(scaled_encoded_data, center = TRUE, scale. = TRUE)
summary(pca_result)
pca_result$rotation
```

`urban` has relatively low absolute loadings compared to other variables, suggesting that it contributes less to the derived components and might not be essential in capturing variance. Then we will go back to try parallel assumption after deleting the `urban` variable.

```{r delete urban}
dass_fit4 =
  dass_fit_df |> 
  mutate(depression_level = factor(case_match(
    depression_level,
    "Normal" ~ "Moderate or Below",
    "Mild" ~ "Moderate or Below","Moderate" ~ "Moderate or Below",
    "Severe" ~ "Severe",
    "Extremely severe" ~ "Extremely severe"
   ), level = c("Moderate or Below", "Severe", "Extremely severe"))) |> 
  polr(depression_level ~ country + religion + race, data = _)

dass_fit4 |>  summary()

dass_fit4 |> poTest()
```
The model meets the parallel assumption now and has only 1 predictor that does not meet parallel assumption.

At this stage, we would like to model the regression line for the selected variables.
```{r}
dass_fit_df$depression_level_numeric <- as.numeric(factor(
  dass_fit_df$depression_level,
  levels = c("Moderate or Below", "Severe", "Extremely severe"),
  labels = c(1, 2, 3)
))

model <- lm(depression_level_numeric ~ country + religion + race - 1, data = dass_fit_df)
summary(model)
```

## Other Variable Selection Methods

Use the model generated by stepwise regression / AIC optimization.

```{r}
dass_fit_step = 
  polr(formula = depression_level ~ country + education + urban + 
    gender + age + religion + orientation + married, data = dass_fit_df)

dass_fit_step |> summary()
```


## Graphs Based On Current Regression Models
```{r}
library(pROC)

# Predicting probabilities for all classes
predicted_probs <- predict(dass_fit_step, type = "probs")

# Extracting the response variable
response <- ifelse(dass_fit_df$depression_level == "Moderate or Below", 1,
                   ifelse(dass_fit_df$depression_level == "Severe", 2, 3))

# Calculating ROC curve for Severe (class 2)
roc_obj <- roc(as.numeric(response == 2), predicted_probs[, "Severe"])

# Plotting ROC curve
plot(roc_obj, col = "blue", main = "ROC Curve for Severe Depression Level")
```

## Future Work

 * Re-selection of independent variables. (Significant, Regroup, AIC, PCA)
 * Interpretation of OR.