---
title: "Logistic Model"
author: "Huanyu Chen, Shaolei Ma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(foreign)
library(ggplot2)
library(MASS)
library(Hmisc)
library(reshape2)
library(car)
library(pROC)
library(effects)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%",
  message = F,
  warning = F
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

The goal is to fit an ordered logit model.

## Independent Variables

First, test the distribution of independent variables.

```{r}
dass_new = read_csv("data/dass_new.csv")

dass_new |> 
  dplyr::select(age, everything()) |> 
  mutate(across(country:depression_level, factor)) |> 
  summary()
```

Some categories contain few samples. For the efficacy of the model, we regroup the independent variables following the standards below:

-   `country`: Group the country with less than 500 samples into "Other", and set "US" to be the reference level.
-   `education`: Order the variable by the length of education.
-   `race`: We follow [standards on race and ethnicity](https://www.govinfo.gov/content/pkg/FR-1997-10-30/pdf/97-28653.pdf) to include only "Asian", "Black", "Native American", "White", and make other categories "Other". We set "White" to be the reference level.
-   `religion`: We conclude "Christian (Catholic)", "Christian (Mormon)", "Christian (Other)", "Christian (Protestant)" into "Christian", and set it as the reference level.
-   `orientation`: Set the "Heterosexual" to be the reference level.
-   `married`: Order the variable by marriage status.

```{r}
dass_fit_df =
  dass_new |> 
  mutate(
    country = relevel(as.factor(ifelse(country %in% c("MY", "US", "GB", "CA", "ID", "PH"), country, "Other")), ref = "US"),
    education = factor(education, levels = c("Less than high school", "High school", "University degree", "Graduate degree")),
    race = relevel(as.factor(ifelse(race %in% c("Arab", "Indigenous Australian"), "Other", race)), ref = "White"),
    religion = relevel(as.factor(ifelse(substr(religion, 1, 9) == "Christian", "Christian", religion)), "Christian"),
    orientation = relevel(as.factor(orientation), "Heterosexual"),
    married = factor(married, levels = c("Never married", "Currently married", "Previously married")),
    depression_level = factor(depression_level, levels = c("Normal", "Mild", "Moderate", "Severe", "Extremely severe"))
  )

dass_fit_df |> summary()
```

## Model Fitting

First fit a model without variables selection.

```{r}
dass_fit = 
  dass_fit_df |> 
  polr(depression_level ~ ., data = _)

(fit_s = 
  dass_fit |> 
  summary())
```

Test the parallel assumption.

```{r}
car::poTest(dass_fit)
```

Because the overall $p-value < 0.001$, we reject the null hypothesis and conclude that the parallel assumption is not met.

To meet the parallel assumption, we try the following two methods:

-   **Select independent variables**: Because we think `religion` and `race`, `married` and `orientation`, `country` and `urban` are strongly correlated with each other, and we care more about `race`, `orientation`, and `country`, first delete `religion`, `married`, and `urban`. We also delete `education` to meet the parallel assumption.
-   **Regroup dependent variable**: Regroup `depression_level` into "Moderate or Below", "Severe", "Extremely severe".

```{r}
dass_fit_df =
  dass_fit_df |> 
  mutate(depression_level = factor(case_match(
    depression_level,
    "Normal" ~ "Moderate or Below",
    "Mild" ~ "Moderate or Below","Moderate" ~ "Moderate or Below",
    "Severe" ~ "Severe",
    "Extremely severe" ~ "Extremely severe"
   ), level = c("Moderate or Below", "Severe", "Extremely severe")))

dass_fit1 =
  dass_fit_df |> 
  polr(depression_level ~ country + gender + age + orientation + race, data = _)

dass_fit1 |> poTest()
```

The model meets the parallel assumption at $\alpha=0.05$ level (with an overall $p = 0.067 > 0.05$).

```{r}
dass_fit1 |> summary()
```

The estimated model can be written as: \begin{align*}
\text{logit}(\hat P(\text{Depression Level} \leq \text{Severe})) &= -0.69 \\
&+ 0.18 \times I(\text{country} = \text{CA}) \\
&+ 0.20 \times I(\text{country} = \text{GB}) \\
&- 0.04 \times I(\text{country} = \text{ID}) \\
&- 0.19 \times I(\text{country} = \text{MY}) \\
&+ 0.13 \times I(\text{country} = \text{PH}) \\
&+ 0.11 \times I(\text{country} = \text{Other}) \\
&- 0.10 \times I(\text{gender} = \text{Male}) \\
&+ 0.38 \times I(\text{gender} = \text{Other}) \\
&- 0.03 \times \text{age} \\
&+ 0.23 \times I(\text{orientation} = \text{Asexual}) \\
&+ 0.40 \times I(\text{orientation} = \text{Bisexual}) \\
&+ 0.30 \times I(\text{orientation} = \text{Homosexual}) \\
&+ 0.22 \times I(\text{orientation} = \text{Other}) \\
&- 0.16 \times I(\text{race} = \text{Asian}) \\
&- 0.09 \times I(\text{race} = \text{Black}) \\
&+ 0.11 \times I(\text{race} = \text{Native American}) \\
&- 0.10 \times I(\text{race} = \text{Other})
\end{align*}

\begin{align*}
\text{logit}(\hat P(\text{Depression Level} \leq \text{Extremely Severe})) &= 0.01 \\
&+ 0.18 \times I(\text{country} = \text{CA}) \\
&+ 0.20 \times I(\text{country} = \text{GB}) \\
&- 0.04 \times I(\text{country} = \text{ID}) \\
&- 0.19 \times I(\text{country} = \text{MY}) \\
&+ 0.13 \times I(\text{country} = \text{PH}) \\
&+ 0.11 \times I(\text{country} = \text{Other}) \\
&- 0.10 \times I(\text{gender} = \text{Male}) \\
&+ 0.38 \times I(\text{gender} = \text{Other}) \\
&- 0.03 \times \text{age} \\
&+ 0.23 \times I(\text{orientation} = \text{Asexual}) \\
&+ 0.40 \times I(\text{orientation} = \text{Bisexual}) \\
&+ 0.30 \times I(\text{orientation} = \text{Homosexual}) \\
&+ 0.22 \times I(\text{orientation} = \text{Other}) \\
&- 0.16 \times I(\text{race} = \text{Asian}) \\
&- 0.09 \times I(\text{race} = \text{Black}) \\
&+ 0.11 \times I(\text{race} = \text{Native American}) \\
&- 0.10 \times I(\text{race} = \text{Other})
\end{align*}

## Effect of Countries on Depression Levels
```{r}
# Extract coefficients for countries
pos_country = str_starts(names(dass_fit1$coefficients), "country")

country_coefficients <- data.frame(
  country = substr(names(dass_fit1$coefficients)[pos_country], 8, 20),
  coefficient = dass_fit1$coefficients[pos_country]
  )

# Plotting coefficients for countries
ggplot(country_coefficients, aes(x = country, y = coefficient)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Effect of Countries on Depression Levels",
       x = "Country", y = "Coefficient")
```

## Effect of Races on Depression Levels
```{r}
# Extract coefficients for races
pos_race <- str_starts(names(dass_fit1$coefficients), "race")

race_coefficients <- data.frame(
  race = substr(names(dass_fit1$coefficients)[pos_race], 5, 20),
  coefficient = dass_fit1$coefficients[pos_race]
)

# Plotting coefficients for races
ggplot(race_coefficients, aes(x = race, y = coefficient)) +
  geom_bar(stat = "identity", fill = "pink", color = "black") +
  labs(
    title = "Effect of Races on Depression Levels",
    x = "Race",
    y = "Coefficient"
  )
```

## Effect of Orientations on Depression Levels
```{r}
# Extract coefficients for orientations
pos_orientation <- str_starts(names(dass_fit1$coefficients), "orientation")

orientation_coefficients <- data.frame(
  orientation = substr(names(dass_fit1$coefficients)[pos_orientation], 12, 20),
  coefficient = dass_fit1$coefficients[pos_orientation]
)

# Plotting coefficients for orientations
ggplot(orientation_coefficients, aes(x = orientation, y = coefficient)) +
  geom_bar(stat = "identity", fill = "palegreen", color = "black") +
  labs(
    title = "Effect of Orientations on Depression Levels",
    x = "Orientation",
    y = "Coefficient"
  )
```

## Interpretation

Calculate the odds ratio and CI for variables:

```{r}
exp(cbind(OR = coef(dass_fit1), confint(dass_fit1))) |> 
  knitr::kable()
```

**Country**:

-   For Canadian, Britain, and Philippines participants, the odds of having a more severe depression level is higher than that of US participants, holding constant all other variables.
-   For Indonesia and Malaysia participants, the odds of having a more severe depression level is lower than that of US participants, holding constant all other variables.

**Gender**:

-   For male participants, the odds of having a more severe depression level is 0.90 times that of female participants, holding constant all other variables.
-   For non-binary participants, the odds of having a more severe depression level is 1.47 times that of female participants, holding constant all other variables.

**Age**:

-   For every one unit increase in participant's age, the odds of having a more severe depression level is multiplied 0.98 times, holding constant all other variables. It indicates that older people may have less risk to get depressed.

**Orientation**:

-   For sexual minorities, the odds of having a more severe depression level is higher than that of heterosexual orientation participants, holding constant all other variables.

**Race**:

-   For Native American participants, the odds of having a more severe depression level is higher than that of White participants, holding constant all other variables.
-   For Asian, Black, and Other participants, the odds of having a more severe depression level is lower than that of White participants, holding constant all other variables.

## Receiver Operating Characteristic Curve for Depression Level
```{r}


# Predicting probabilities for all classes
predicted_probs <- predict(dass_fit1, type = "probs")

# Extracting the response variable
response <- ifelse(dass_fit_df$depression_level == "Moderate or Below", 1,
                   ifelse(dass_fit_df$depression_level == "Severe", 2, 3))

# Calculating ROC curves for three responses
roc_obj_1 <- roc(as.numeric(response == 1), predicted_probs[, "Moderate or Below"])
roc_obj_2 <- roc(as.numeric(response == 2), predicted_probs[, "Severe"])
roc_obj_3 <- roc(as.numeric(response == 3), predicted_probs[, "Extremely severe"])

# Plotting ROC curves for three responses on the same graph
plot(roc_obj_1, col = "blue", main = "ROC Curves for Three Responses")
plot(roc_obj_2, col = "red", add = TRUE)
plot(roc_obj_3, col = "green", add = TRUE)
legend("bottomright",
       legend = c("Moderate or Below", "Severe", "Extremely Severe"),
       col = c("blue", "red", "green"), lty = 1)
```
According to this graph, the ROC curves for "Moderate or Below" and "Extremely Severe" are close to the top-left corner of the plot, showing a good model's performance in these two categories. The performance on "Severe" may indicate that we need more data points that have a depression level of "Severe."

## Effects of Predictors on Depression Levels
The following graphs show the probability of depression level on different categorical variables.
```{r, fig.width=20, fig.height=50, message = FALSE}
# Create a data frame with the predictor values for visualization
predictor_data <- model.matrix(depression_level ~ country + gender + age
                               + orientation + race, data = dass_fit_df)

# Compute the effects
effects <- allEffects(dass_fit1, xlevels = list(predictor_data))

# Plot the effects
plot(effects, style = "stacked")
```

## Limitations

-   Imbalance within samples (e.g. 76.3% female samples).
-   Strong dependence between independent variables (e.g. race versus religion).
-   Counterintuitive results (e.g. Asian people have lower risk of getting severely depressed than white people).
-   Future Improvements:
    -   Clustering to decrease dependence;
    -   Fit a separate model for each depression level to avoid parallel assumption.
