---
title: "Depression among the World"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)

library(plotly)
library(countrycode)
library(shiny)

dass_df = read_csv("data/dass_new.csv")

country_df =
  dass_df |> 
  count(country, depression_level) |> 
  group_by(country) |> 
  mutate(per =  n / sum(n),
         code = ifelse(country == "XK", "XKX", countrycode(country, origin = "iso2c", destination = "iso3c")),
         country = ifelse(country == "XK", "Kosovo", countrycode(country, origin = "iso2c", destination = "country.name"))
  )

```

## Column {.sidebar}

```{r}
dlevel =
  dass_df |>
  distinct(depression_level) |>
  pull()

# selectInput widget
selectInput(
  inputId = "Depression Level", 
  label = h3("Select Level"),
  choices = dlevel,
  selected = "Normal")
```

## Row {data-width="650"}

```{r}
renderPlotly({
  # light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

country_df |> 
  filter(depression_level == input[["Depression Level"]]) |> 
  plot_geo() |> 
  add_trace(
  z = ~per, color = ~per, colors = 'Blues',
  text = ~ country, locations = ~ code, marker = list(line = l)
  ) |> 
  colorbar(title = "Percentage") |> 
  layout(
    title = str_c("Percentage of ",input[["Depression Level"]], " depression among each country"),
    geo = g
  )
})
```

## Row {.tabset .tabset-fade}
### Description
This map shows how different depression levels spread among the world.

\ \par

For example, if selecting 'normal' level (also default level), the plot would show the percentage of responses with normal depression level for each country around the world.

### Formula

For each country, we can calculate:

percentage = number of responses at the selected level / number of total responses

