---
title: "Depression among the World"
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 3
      bootswatch: flatly
    orientation: rows
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)

library(plotly)
library(countrycode)
library(shiny)

dass_df = read_csv("data/dass_new.csv")

country_df =
  dass_df |> 
  count(country, depression_level) |> 
  group_by(country) |> 
  mutate(
    proportion =  n / sum(n),
    code = ifelse(country == "XK", "XKX", countrycode(country, origin = "iso2c", destination = "iso3c")),
    country = ifelse(country == "XK", "Kosovo", countrycode(country, origin = "iso2c", destination = "country.name")),
    depression_level = factor(depression_level, levels = c("Normal", "Mild", "Moderate", "Severe", "Extremely severe"), ordered = T)
  ) |> 
  rename(occurrences = n)
```

## Column {.sidebar}

```{r}
dlevel =
  dass_df |>
  distinct(depression_level) |>
  pull()

# selectInput widget
selectInput(
  inputId = "Depression Level",
  label = h3("Select Depression Level"),
  choices = dlevel,
  selected = "Normal")

selectInput(
  inputId = "Value Type",
  label = h3("Occurrences / Proportion"),
  choices = c("Occurrences", "Proportion"),
  selected = "Normal")

sliderInput(
  inputId = "Occurrences Range",
  label = h3("Select occurrences range"),
  min = 0, max = 6000,
  value = c(100, 6000)
)
```

 
## Row {data-width="750"}

```{r}
renderPlotly({
  # light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

country_df |> 
  filter(depression_level == input[["Depression Level"]]) |> 
  group_by(country, code) |> 
  summarise(
    occurrences = sum(occurrences),
    proportion = sum(proportion)
  ) |>
  filter(
    occurrences >= input[["Occurrences Range"]][1],
    occurrences <= input[["Occurrences Range"]][2]
  ) |> 
  select(country, code, val = tolower(input[["Value Type"]])) |> 
  plot_geo() |> 
  add_trace(
  z = ~ val, color = ~ val, colors = 'Blues',
  text = ~ country, locations = ~ code, marker = list(line = l)
  ) |> 
  colorbar(title = input[["Value Type"]]) |> 
  layout(
    title = str_c(input[["Value Type"]], " of ",input[["Depression Level"]], " Depression Among Each Country"),
    geo = g
  )
})
```

## Row {.tabset .tabset-fade}
### Description
This map shows how different depression levels spread among the world.

\ \par

For example, if selecting "Normal" level (also default level), "Occurrences", and the default occurrences range "100 - 6000", the plot would show the occurrences of responses with normal depression level for countries with 100 - 6000 normal depression level responses.

\ \par

It could be observed that Malaysia and United States have the largest number of responses. Among all the depression levels, extremely severe depression takes up the largest proportion among most countries.

However, the interactive graph also illustrates the imbalance of responses. If we limit the occurrences to be larger than 500, there will only be Malaysia and United States left. When interpreting the proportion of different depression level, the sample size must be considered.

### Formula

For each country, we can calculate:

\ \par

percentage = number of responses at the selected level / number of total responses

